Profiling Walksheds

### 2023-06-21 data-raw/02-OSM-Amenities.Rmd
Data objects created
NECTAR presentation

---
title: "Network-Preprocessing"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: pdf_document
---

@@

Obtain the coordinates of the points:
```{r}
amenities_coords <- amenities_points |>
  st_coordinates()

amenities_points <- amenities_points |>
  mutate(lon = amenities_coords[,1], lat = amenities_coords[,2])
```

Use the coordinates to obtain the h3 indices at level 13:
```{r}
amenities_points <- amenities_points |>
  mutate(h3_index = geo_to_h3(st_coordinates(geometry),
                              13))
```

>>>>

---
title: "OSM Extracting Amenities"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: html_notebook
---

**NOTE:** Notebook takes minutes to run.

@@

Obtain the h3 indices at level 13 from the simple features object:
```{r}
amenities_h3 <- geo_to_h3(amenities_points, res = 13)
```

Add h3 indices to amenities_points data object:
```{r}
amenities_points$h3_index <- amenities_h3
```

### 2023-06-21 data-raw/03-Network-Attributes-by-Walkshed.Rmd
NECTAR presentation

@@

In this document I aim to partition the network based on walksheds.

@@

for(i in 1:nrow(walksheds_da)){
  wsnet <- st_intersection(hamilton_edges, 
                           walksheds_da[i,]) |> 
    st_cast(to = "LINESTRING")
  wsnet <- as_sfnetwork(wsnet)
  E(wsnet)$weight <- E(wsnet)$length
  walksheds_da_net_vars$n_nodes[i] <- gorder(wsnet)
  walksheds_da_net_vars$transitivity[i] <- transitivity(wsnet)
  walksheds_da_net_vars$motifs_3[i] <- count_motifs(wsnet, size = 3)
  walksheds_da_net_vars$motifs_4[i] <- count_motifs(wsnet, size = 4)
  walksheds_da_net_vars$n_edges[i] <- gsize(wsnet)
  walksheds_da_net_vars$diameter[i] <- diameter(wsnet)
  walksheds_da_net_vars$radius[i] <- radius(wsnet)
  walksheds_da_net_vars$girth[i] <- girth(wsnet)$girth
  walksheds_da_net_vars$global_efficiency[i] <- global_efficiency(wsnet)
  walksheds_da_net_vars$edge_density[i] <- edge_density(wsnet)
  walksheds_da_net_vars$edge_connectivity[i] <- edge_connectivity(wsnet)
  walksheds_da_net_vars$mean_distance[i] <- mean_distance(wsnet)
  walksheds_da_net_vars$min_cut[i] <- min_cut(wsnet)
}

@@

for(i in 1:nrow(walksheds_validation)){
  wsnet <- st_intersection(hamilton_edges, 
                           walksheds_validation[i,]) |> 
    st_cast(to = "LINESTRING")
  wsnet <- as_sfnetwork(wsnet)
  E(wsnet)$weight <- E(wsnet)$length
  walksheds_validation_net_vars$n_nodes[i] <- gorder(wsnet)
  walksheds_validation_net_vars$transitivity[i] <- transitivity(wsnet)
  walksheds_validation_net_vars$motifs_3[i] <- count_motifs(wsnet, size = 3)
  walksheds_validation_net_vars$motifs_4[i] <- count_motifs(wsnet, size = 4)
  walksheds_validation_net_vars$n_edges[i] <- gsize(wsnet)
  walksheds_validation_net_vars$diameter[i] <- diameter(wsnet)
  walksheds_validation_net_vars$radius[i] <- radius(wsnet)
  walksheds_validation_net_vars$girth[i] <- girth(wsnet)$girth
  walksheds_validation_net_vars$global_efficiency[i] <- global_efficiency(wsnet)
  walksheds_validation_net_vars$edge_density[i] <- edge_density(wsnet)
  walksheds_validation_net_vars$edge_connectivity[i] <- edge_connectivity(wsnet)
  walksheds_validation_net_vars$mean_distance[i] <- mean_distance(wsnet)
  walksheds_validation_net_vars$min_cut[i] <- min_cut(wsnet)
}

@@

>>>>

**NOTE:** Notebook takes about 7 hours to run.

@@

In this document we aim to partition the network based on walksheds.

We tried smoothing the pseudo-nodes, but this process messed up with the geometry of the network which affects some of the network analysis, so we decided not to proceed. Instead, we use the network as is. Intersecting the network mutilates all edges that cross the boundary of the pedshed. To avoid this we keep track of the indices of the edges that intersect the pedshed and then retrieve them from the network.

@@

for(i in 1:nrow(walksheds_da)){
  # Find the intersection between the edges and the walksheds. The edges will be truncated, but the indices can be used to retrieve the full edges.
  wsnet <- hamilton_edges |>
    st_filter(
      walksheds_da[i,],
      .predicate = st_intersects)

  # Extract the full edges that were found to intersect the walkshed
  wsnet <- hamilton_edges |> 
    filter(edge_index %in% wsnet$edge_index)

  # Convert to sfnetwork
  wsnet <- as_sfnetwork(wsnet)

  E(wsnet)$weight <- E(wsnet)$length
  walksheds_da_net_vars$n_nodes[i] <- gorder(wsnet)
  walksheds_da_net_vars$transitivity[i] <- transitivity(wsnet)
  walksheds_da_net_vars$motifs_3[i] <- count_motifs(wsnet, size = 3)
  walksheds_da_net_vars$motifs_4[i] <- count_motifs(wsnet, size = 4)
  walksheds_da_net_vars$n_edges[i] <- gsize(wsnet)
  walksheds_da_net_vars$diameter[i] <- diameter(wsnet)
  walksheds_da_net_vars$radius[i] <- radius(wsnet)
  walksheds_da_net_vars$girth[i] <- girth(wsnet)$girth
  walksheds_da_net_vars$global_efficiency[i] <- global_efficiency(wsnet)
  walksheds_da_net_vars$edge_density[i] <- edge_density(wsnet)
  walksheds_da_net_vars$edge_connectivity[i] <- edge_connectivity(wsnet)
  walksheds_da_net_vars$mean_distance[i] <- mean_distance(wsnet)
  walksheds_da_net_vars$min_cut[i] <- min_cut(wsnet)
}

@@

for(i in 1:nrow(walksheds_validation)){
  # Find the intersection between the edges and the walksheds. The edges will be truncated, but the indices can be used to retrieve the full edges.
  wsnet <- hamilton_edges |>
    st_filter(
      walksheds_validation[i,],
      .predicate = st_intersects)

  # Extract the full edges that were found to intersect the walkshed
  wsnet <- hamilton_edges |> 
    filter(edge_index %in% wsnet$edge_index)

  # Convert to sfnetwork
  wsnet <- as_sfnetwork(wsnet)

  E(wsnet)$weight <- E(wsnet)$length
  walksheds_validation_net_vars$n_nodes[i] <- gorder(wsnet)
  walksheds_validation_net_vars$transitivity[i] <- transitivity(wsnet)
  walksheds_validation_net_vars$motifs_3[i] <- count_motifs(wsnet, size = 3)
  walksheds_validation_net_vars$motifs_4[i] <- count_motifs(wsnet, size = 4)
  walksheds_validation_net_vars$n_edges[i] <- gsize(wsnet)
  walksheds_validation_net_vars$diameter[i] <- diameter(wsnet)
  walksheds_validation_net_vars$radius[i] <- radius(wsnet)
  walksheds_validation_net_vars$girth[i] <- girth(wsnet)$girth
  walksheds_validation_net_vars$global_efficiency[i] <- global_efficiency(wsnet)
  walksheds_validation_net_vars$edge_density[i] <- edge_density(wsnet)
  walksheds_validation_net_vars$edge_connectivity[i] <- edge_connectivity(wsnet)
  walksheds_validation_net_vars$mean_distance[i] <- mean_distance(wsnet)
  walksheds_validation_net_vars$min_cut[i] <- min_cut(wsnet)
}

@@

Example of pedshed with network:
```{r}
sample_walkshed <- sample.int(nrow(walksheds_da), 1)

wsnet <- hamilton_edges |>
  st_filter(
    walksheds_da[sample_walkshed,],
    .predicate = st_intersects)

# Extract the full edges that were found to intersect the walkshed
wsnet <- hamilton_edges |> 
  filter(edge_index %in% wsnet$edge_index)

ggplot() +
  geom_sf(data = walksheds_da[sample_walkshed,]) +
  geom_sf(data = wsnet) +
  theme_void()

### 2023-06-21 data-raw/04-Accessibility-Calculations.Rmd
NECTAR presentation

@@

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(tidyr) # Tidy Messy Data
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r}
#load(file = "hamilton_net.rda")
load(file = "amenities_points.rda")
load(file = "urban_hammer_da.rda")
data(urban_types)
load(file = "walksheds.rda")
```

@@

amenities_walksheds <- amenities_points |>
  select(amenity, Category) |>
  st_intersection(walksheds)

@@

Save amenities by walkshed:
```{r}
save(amenities_walksheds, file = "amenities_walkshed.rda", compress = "xz")
```

Summary of amenity categories by DA and Type:
```{r}
amenities_walksheds_summary <- amenities_walksheds |>
  drop_na() |>
  st_drop_geometry() |>
  group_by(GeoUID,
           Type,
           Category) |>
  summarize(n = n(),
            .groups = "drop") |>
  pivot_wider(id_cols = c(GeoUID, Type),
              names_from = Category,
              values_from = n,
              values_fill = 0)
```

@@

Almost 50% of suburban population (which represent about 23% of the total population in Hamilton) live in DAs with zero amenities within the 15-min walk shed.

About 18% of urban population (representing about 5% of the total population of the city) live in DAs with zero amenities within the 15-min walk shed.

@@

Barely 2.5% of suburban residents live in DAs with minimal 15-min walk sheds. They represent approximately half a percentage point of the total population of the city.

More than 15% of urban population live in DAs with minimal 15-min walk sheds, and account for approximately 2.6% of the total population of the city.

@@

# 409029, 534918

sample_walksheds <- amenities_walksheds |>
  group_by(Type) |>
  slice_sample(n = 1)

ggplot()  + 
  # geom_sf(data = urban_types |>
  #               filter(Type != "Rural"),
  #         aes(fill = Type)) +
  geom_sf(data = walksheds |>
            filter((GeoUID == sample_walksheds$GeoUID[1] & Type == sample_walksheds$Type[1]) | 
                     (GeoUID == sample_walksheds$GeoUID[2] & Type == sample_walksheds$Type[2])),
          aes(fill = Type),
          color = "black")+
  geom_sf(data = amenities_walksheds |>
            filter((GeoUID == sample_walksheds$GeoUID[1] & Type == sample_walksheds$Type[1]) | 
                     (GeoUID == sample_walksheds$GeoUID[2] & Type == sample_walksheds$Type[2])),
          #aes(color = Category)
          )
```

>>>>

**NOTE:** Notebook takes about 10 minutes to run.

@@

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(h3)
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(tidyr) # Tidy Messy Data
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r}
#load(file = "hamilton_net.rda")
load(file = "amenities_points.rda")
load(file = "urban_hammer_da.rda")
data(urban_types)
load(file = "walksheds_da.rda")
load(file = "walksheds_validation.rda")
```

@@

amenities_walksheds_da <- amenities_points |>
  select(amenity, Category) |>
  st_intersection(walksheds_da)

@@

Validation walksheds:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

amenities_walksheds_validation <- amenities_points |>
  select(amenity, Category) |>
  st_intersection(walksheds_validation)

Save amenities by walkshed:
```{r}
save(amenities_walksheds_da, file = "amenities_walkshed_da.rda", compress = "xz")
save(amenities_walksheds_validation, file = "amenities_walkshed_validation.rda", compress = "xz")
```

Summary of amenity categories by DA and Type:
```{r}
amenities_walksheds_summary <- amenities_walksheds_da |>
  drop_na() |>
  st_drop_geometry() |>
  group_by(GeoUID,
           Type,
           Category) |>
  summarize(n = n(),
            .groups = "drop") |>
  pivot_wider(id_cols = c(GeoUID, Type),
              names_from = Category,
              values_from = n,
              values_fill = 0)
```

@@

About 50% of suburban population (which represent about 22% of the total population in Hamilton) live in DAs with zero amenities within the 15-min walk shed.

About 15% of urban population (representing about 3.7% of the total population of the city) live in DAs with zero amenities within the 15-min walk shed.

@@

Barely 2.3% of suburban residents live in DAs with minimal 15-min walk sheds. They represent approximately 1% of the total population of the city.

More than 15% of urban population live in DAs with minimal 15-min walk sheds, and account for approximately 3.8% of the total population of the city.

@@

# 609054, 144675, 741580

sample_walksheds <- amenities_walksheds_da |>
  group_by(Type) |>
  slice_sample(n = 1)

ggplot()  + 
  # geom_sf(data = urban_types |>
  #               filter(Type != "Rural"),
  #         aes(fill = Type)) +
  geom_sf(data = walksheds_da |>
            filter((GeoUID == sample_walksheds$GeoUID[1] & Type == sample_walksheds$Type[1]) | 
                     (GeoUID == sample_walksheds$GeoUID[2] & Type == sample_walksheds$Type[2])),
          aes(fill = Type),
          color = "black")+
  geom_sf(data = amenities_walksheds_da |>
            filter((GeoUID == sample_walksheds$GeoUID[1] & Type == sample_walksheds$Type[1]) | 
                     (GeoUID == sample_walksheds$GeoUID[2] & Type == sample_walksheds$Type[2])),
          #aes(color = Category)
          )
```

Distance to amenities:
```{r}
load("ttm_walk_da.rda")
```

```{r}
amenities_by_h3 <- amenities_points |>
  st_drop_geometry() |>
  group_by(h3_index) |>
  summarize(n_amenities = n(),
            .groups = "drop")
```

Join the amenities to the travel time matrix by means of the h3 indices:
```{r}
ttm_walk_to_amenities <- ttm_walk_da |>
  left_join(amenities_by_h3,
            by = "h3_index") |>
  drop_na(n_amenities)
```

Also join the urban fabric type:
```{r}
ttm_walk_to_amenities <- ttm_walk_to_amenities |>
  left_join(urban_hammer_da |>
              st_drop_geometry() |>
              transmute(GeoUID, Type = Type.1),
            by = "GeoUID")
```

Summarize travel times to amenities from DA centroids:
```{r}
junk <- ttm_walk_to_amenities |>
  select(GeoUID, Type, travel_time) |>
  group_by(GeoUID, Type) |>
  summarize(min_travel_time = min(travel_time),
            mean_travel_time = mean(travel_time),
            .groups = "drop")
```

Distribution of mean travel times by urban fabric:
```{r}
junk |>
  ggplot(aes(x = mean_travel_time,
             color = Type)) +
  geom_density()
```

Distribution of min travel times by urban fabric:
```{r}
junk |>
  ggplot(aes(x = min_travel_time,
             color = Type)) +
  geom_density()
```

### 2023-06-21 data-raw/05-Walkshed-Profiles.Rmd
NECTAR presentation

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(patchwork) # The Composer of Plots
library(tidyr) # Tidy Messy Data
library(tree) # Classification and Regression Trees
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data:
```{r}
load(file = "amenities_walkshed.rda")
load(file = "hamilton_net.rda")
load(file = "urban_hammer_da.rda")
data("urban_types")
load(file = "walksheds.rda")
load(file = "walksheds_net_attributes.rda")
```

@@

The aim is to train a classification tree to profile urban and suburban walksheds based on the attributes of the network. We need to drop 1 observation that has an NA in `transitivity`:
```{r}
d <- walksheds_net_attributes |>
                  drop_na()
```

@@

Tree 1 with seed 1 (912571):
```{r}
#seed1 <- sample.int(1000000, 1)
seed1 <- 912571
set.seed(seed1)
tree1 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree1)
```

Tree 2 with seed 2 (94766):
```{r}
#seed2 <- sample.int(1000000, 1)
seed2 <- 94766
set.seed(seed2)
tree2 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree2)
```

Tree 3 with seed 3 (46125):
```{r}
#seed3 <- sample.int(1000000, 1)
seed3 <- 46125
set.seed(seed3)
tree3 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree3)
```

Tree 4 with seed 4 (556090)
```{r}
#seed4 <- sample.int(1000000, 1)
seed4 <- 556090
set.seed(seed4)
tree4 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree4)
```

```{r}
#seed4 <- sample.int(1000000, 1)
seed4 <- 556090
set.seed(seed4)
tree5 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
summary(tree5)
```

@@

Notice that transitivity and motifs are the dominant attributes in the classification trees. Trees 1, 3, and 4 have the same general structure, and in fact trees 3 and 4 are identical. Based on this we choose to use `tree3`.

@@

>>>>

Load packages:
```{r warning=FALSE, message=FALSE}
library(broom) # Convert Statistical Objects into Tidy Tibbles
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(gt) # Easily Create Presentation-Ready Display Tables
library(patchwork) # The Composer of Plots
library(tidyr) # Tidy Messy Data
library(tree) # Classification and Regression Trees
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data:
```{r}
load(file = "amenities_walkshed_da.rda")
load(file = "hamilton_net.rda")
load(file = "urban_hammer_da.rda")
data("urban_types")
load(file = "walksheds_da.rda")
load(file = "walksheds_da_net_vars.rda")
```

@@

The aim is to train a classification tree to profile urban and suburban walksheds based on the attributes of the network. We need to drop 1 observation that has an NA in `transitivity`:
```{r}
d <- walksheds_da_net_vars |>
                  drop_na() |>
  mutate(normalized_motifs_3 = motifs_3/n_edges,
         normalized_motifs_4 = motifs_4/n_edges)
```

@@

Tree 1 with seed 1 (912571):
```{r}
seed1 <- sample.int(1000000, 1)
#seed1 <- 908544
set.seed(908544)
tree1 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree1)
```

Tree 2 with seed 2 (94766):
```{r}
seed2 <- sample.int(1000000, 1)
#seed2 <- 669402
seed2
set.seed(seed2)
tree2 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree2)
```

Tree 3 with seed 3 (46125):
```{r}
seed3 <- sample.int(1000000, 1)
#seed3 <- 853251
seed3
set.seed(seed3)
tree3 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree3)
```

Tree 4 with seed 4 (556090)
```{r}
seed4 <- sample.int(1000000, 1)
#seed4 <- 707633
seed4
set.seed(seed4)
tree4 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree4)
```

@@

Notice that edge_density and normalized_motifs_3 are the dominant attributes in the classification trees. Trees 1, 3, and 4 have the same general structure, and in fact trees 3 and 4 are identical. Based on this we choose to use `tree3`.

@@

```{r add-classification-to-data}
#| include: false

d <- d |>
  mutate(leaf = case_when(edge_density < 0.00591859212548868 & normalized_motifs_3 < 0.86734693877551 ~ "Leaf 1",
                          edge_density < 0.00591859212548868 & normalized_motifs_3 >= 0.86734693877551 ~ "Leaf 2",
                          edge_density >= 0.00591859212548868 & normalized_motifs_3 < 0.967741935483871 ~ "Leaf 3",
                          edge_density >= 0.00591859212548868 & normalized_motifs_3 >= 0.967741935483871 ~ "Leaf 4"),
         leaf = factor(leaf))
```

```{r join-area-of-ped-sheds}
#| include: false

d <- d |>
  left_join(walksheds |>
              mutate(area = st_area(geometry),
                     area = units::set_units(area, km^2) |> 
                       drop_units()) |>
              st_drop_geometry(),
            by = c("GeoUID", "Type")) 
```

```{r area-of-ped-sheds-by-classification}
#| echo: false

d |>
  select(area, leaf) |>
  group_by(leaf) |>
  skim() |>
  mutate(across(numeric.mean:numeric.p100, 
                      ~round(.x, 2)),
         Type = c("Typically suburban", "Typically urban", "Typically suburban", "Typically urban")) |>
  select(leaf, Type, starts_with("numeric")) |>
  rename_with(~ str_replace(.x, "numeric.", ""), starts_with("numeric")) |>
  gt() |>
  cols_width(everything() ~ px(100)) |>
  tab_options(table.font.size = 30)
```

Typically suburban networks contain many unique patterns (few motifs).

0.00591859212548868 & normalized_motifs_3 < 0.86734693877551 ~ "Leaf 1",
                          edge_density < 0.00591859212548868 & normalized_motifs_3 >= 0.86734693877551 ~ "Leaf 2",
                          edge_density >= 0.00591859212548868 & normalized_motifs_3 < 0.967741935483871 ~ "Leaf 3",

```{r plot-leaf-1}
set.seed(29919)

leaf_1_plot <- st_intersection(hamilton_net$edges,
             walksheds_da |> 
  filter(GeoUID == (walksheds_da_net_vars |> 
                      mutate(normalized_motifs_3 = motifs_3/n_edges) |>
                      filter(edge_density < 0.00591859212548868,
                             normalized_motifs_3 < 0.86734693877551, 
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID))))

leaf_1_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_1_plot$edge_index)

#leaf_1_plot <- 
  leaf_1_plot |>
  ggplot() +
  geom_sf() +
  ggtitle("edge density < 0.006, motifs < 0.867") +
  theme_void()
```

```{r plot-leaf-2}
set.seed(84451)

leaf_2_plot <- st_intersection(hamilton_net$edges,
             walksheds_da |> 
  filter(GeoUID == (walksheds_da_net_vars |> 
                      mutate(normalized_motifs_3 = motifs_3/n_edges) |>
                      filter(edge_density >= 0.00591859212548868,
                             normalized_motifs_3 < 0.967741935483871, 
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID))))

leaf_2_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_2_plot$edge_index)

leaf_2_plot |>
  ggplot() +
  geom_sf() +
  ggtitle("edge density < 0.006, motifs >= 0.968") +
  theme_void()
```

```{r display-suburban-networks}
#| echo: false

leaf_1_plot + plot_spacer() + leaf_3_plot
```

Typically urban networks contain isomorphic patterns that repeat (large number of motifs).

```{r plot-leaf-3}
set.seed(92348)

leaf_1_plot <- st_intersection(hamilton_net$edges,
             walksheds_da |> 
  filter(GeoUID == (walksheds_da_net_vars |> 
                      mutate(normalized_motifs_3 = motifs_3/n_edges) |>
                      filter(edge_density < 0.00591859212548868,
                             normalized_motifs_3 >= 0.86734693877551, 
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID))))

leaf_1_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_1_plot$edge_index)

#leaf_1_plot <- 
  leaf_1_plot |>
  ggplot() +
  geom_sf() +
  ggtitle("edge density < 0.006, motifs >= 0.867") +
  theme_void()
```

```{r plot-leaf-4}
set.seed(392847)

leaf_4_plot <- st_intersection(hamilton_net$edges,
             walksheds_da |> 
  filter(GeoUID == (walksheds_da_net_vars |> 
                      mutate(normalized_motifs_3 = motifs_3/n_edges) |>
                      filter(edge_density >= 0.00591859212548868,
                             normalized_motifs_3 >= 0.967741935483871, 
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID))))

leaf_4_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_4_plot$edge_index)

leaf_4_plot |>
  ggplot() +
  geom_sf() +
  ggtitle("edge density < 0.006, motifs >= 0.968") +
  theme_void()
```

```{r display-urban-networks}
#| echo: false

leaf_2_plot + plot_spacer() + leaf_4_plot
```

## Analysis of pedshed areas as a function of network attributes

```{r estimate-model-area}
#| include: false
mdl <- d |>
  mutate(area = log(area),
         edge_density_2 = edge_density^2,
         normalized_motifs_3_2 = normalized_motifs_3^2) |>
  lm(area ~ edge_density + edge_density_2 + normalized_motifs_3, data = _)
```

```{r tidy-results-of-model-area}
#| include: false

mdl_tidy <- mdl |>
  tidy() |>
  mutate(term = c("intercept", "edge_density", "edge_density^2", "normalized_motifs_3"))
```

```{r plot-effects-transitivity}
#| include: false

b1 <- mdl_tidy |> filter(term == "intercept") |> pull(estimate)
b2 <- mdl_tidy |> filter(term == "edge_density") |> pull(estimate)
b3 <- mdl_tidy |> filter(term == "edge_density^2") |> pull(estimate)
f_transitivity <- function(x) exp(b1 + b2 * x +  b3 * x^2)

effect_transitivity <- ggplot() +
  geom_function(fun = f_transitivity,
                xlim = c(0, 0.4)) +
    ylim(c(0.05, 1.9)) +
  ggtitle("Expected area as a function of transitivity") + 
  ylab("area (sq.km)") +
  xlab("transitivity") +
  theme_minimal()
```

```{r plot-effect-motifs}
#| include: false

b1 <- mdl_tidy |> filter(term == "intercept") |> pull(estimate)
b4 <- mdl_tidy |> filter(term == "normalized_motifs_3") |> pull(estimate)
f_motifs <- function(x) exp(b1 + b4 * x)

effect_motifs <- ggplot() +
  geom_function(fun = f_motifs,
                xlim = c(0.45, 1.1)) +
  ylim(c(0.05, 1.9)) +
  ggtitle("Expected area as a function of normalized motifs (size = 3)") + 
  ylab("area (sq.km)") +
  xlab("normalized motifs (size = 3)") +
  theme_minimal()
```

```{r display-model-effects}
#| echo: false

effect_transitivity + effect_motifs
```

<!--
Check the break labels of `tree3` to round:
```{r}
ggparty(tree3)$data$breaks_label
```
